\documentclass[10pt,a4paper]{article}
\usepackage[margin=0.62in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,calc,shapes.geometric,decorations.pathreplacing,fit}

\definecolor{brand}{HTML}{0D3B66}
\definecolor{accent}{HTML}{2A9D8F}
\definecolor{softbg}{HTML}{F4F8FB}
\definecolor{textdark}{HTML}{1E1E1E}

\hypersetup{
  colorlinks=true,
  linkcolor=brand,
  urlcolor=accent
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{3pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=1pt,topsep=2pt}

\titleformat{\section}{\color{brand}\large\bfseries}{\thesection.}{0.4em}{}
\titleformat{\subsection}{\color{textdark}\bfseries}{\thesubsection}{0.4em}{}
\titlespacing*{\section}{0pt}{8pt}{4pt}
\titlespacing*{\subsection}{0pt}{6pt}{3pt}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small CS671 Technical Model Report}
\fancyhead[R]{\small Paired Diffusion (Pelvis)}
\fancyfoot[C]{\small \thepage}
\setlength{\headheight}{14pt}

\begin{document}

% -------------------- Header Block --------------------
\begin{tcolorbox}[
  enhanced,
  colback=softbg,
  colframe=brand,
  boxrule=0.9pt,
  arc=2mm,
  left=5pt,right=5pt,top=5pt,bottom=5pt
]
\centering
{\Large\bfseries\color{brand} CS671 -- Deep Learning and Its Applications}\\[2pt]
{\large\bfseries Technical Documentation Report}\\[6pt]
{\normalsize\bfseries Model: Paired Diffusion (Pelvis)}\\[4pt]
\small
Project Group: 5 \quad Region: Pelvis \quad Modality: Paired (CT $\to$ MRI)
\end{tcolorbox}

\section{ARCHITECTURAL OVERVIEW}
Paired Diffusion utilizes a Latent Diffusion Model (LDM) framework to perform CT-to-MRI synthesis. Unlike GAN-based approaches, diffusion models learn the data distribution by reversing a gradual denoising process. In this implementation, the model is conditioned on CT features to guide the generation of aligned MRI slices in a compressed latent space.

\subsection{Latent Diffusion Architecture}
The implementation utilizes a high-fidelity Latent Diffusion Model (LDM) where denoising occurs in a compressed $64 \times 64 \times 4$ latent space. This projet employs a modified Variational Autoencoder (VAE) with custom skip-connections to ensure the preservation of complex pelvic anatomical structures.

\begin{figure}[H]
\centering
\resizebox{\textwidth}{!}{
\begin{tikzpicture}[
    node distance=5mm and 10mm,
    block/.style={draw=brand, thick, fill=softbg, minimum width=2.4cm, minimum height=0.9cm, align=center, rounded corners=1mm, font=\scriptsize},
    module/.style={draw=accent, thick, fill=white, minimum width=2.0cm, minimum height=0.7cm, align=center, font=\tiny},
    arr/.style={-{Latex[length=1.5mm]}, thick, draw=brand!70},
    skip/.style={-{Latex[length=1.5mm]}, thick, dashed, draw=accent}
]
    % VAE Encoder Path
    \node[block] (in_mri) {Real MRI\\(y)};
    \node[module, right=of in_mri] (enc_1) {VAE Encoder\\Block 1};
    \node[module, right=of enc_1] (enc_2) {VAE Encoder\\Block 2};
    \node[block, right=of enc_2, fill=accent!10] (latent) {Latent $z$\\ (4, 64, 64)};

    % Denoising Path (Middle)
    \node[block, below=1.5cm of enc_1] (ct_in) {Input CT\\(x)};
    \node[module, right=of ct_in] (ct_enc) {CT Encoder};
    \node[draw=brand, thick, dashed, rounded corners=2mm, inner sep=3mm, right=3.5cm of ct_enc] (unet) {Denoising UNet\\($\epsilon_\theta$)};
    
    % VAE Decoder Path (Bottom)
    \node[block, right=2.0cm of unet, fill=accent!10] (z_hat) {Denoised $\hat{z}$};
    \node[module, right=of z_hat] (dec_1) {VAE Decoder\\Block 1};
    \node[module, right=of dec_1] (dec_2) {VAE Decoder\\Block 2};
    \node[block, right=of dec_2] (out_mri) {Synthetic MRI\\($\hat{y}$)};

    % Connections
    \draw[arr] (in_mri) -- (enc_1); \draw[arr] (enc_1) -- (enc_2); \draw[arr] (enc_2) -- (latent);
    \draw[arr] (latent) -- (unet) node[midway, right=2mm, font=\tiny] {+ Noise};
    \draw[arr] (ct_in) -- (ct_enc);
    \draw[arr] (ct_enc) -- (unet);
    \draw[arr] (unet) -- (z_hat);
    \draw[arr] (z_hat) -- (dec_1); \draw[arr] (dec_1) -- (dec_2); \draw[arr] (dec_2) -- (out_mri);

    % Skip Connections
    \draw[skip] (enc_1.south) |- ($(dec_2.north) + (0,0.5)$) -- (dec_2.north) node[pos=0.8, right, font=\tiny, color=accent] {skip\_conv};
    
    % Grouping
    \node[draw=accent, dashed, thick, rounded corners=2mm, inner sep=2mm, fit=(ct_enc) (unet), label={[font=\tiny\bfseries, color=accent]below:Latent Denoising Pipeline}] (group) {};
\end{tikzpicture}
}
\caption{Enhanced Latent Diffusion Architecture: Implementation-specific VAE skip-pathways project intermediate encoder activations into the decoder to maintain high-resolution organ boundaries (prostate, bladder) during MRI synthesis.}
\end{figure}

\textbf{VAE Skip-Connection Injection:} 
The \texttt{VAEDecode} implementation manually injects captured activations from the encoder's \texttt{down\_blocks} into the decoder's \texttt{up\_blocks}. This is achieved via four $1 \times 1$ convolution layers (\texttt{skip\_conv\_1-4}) that align the feature dimensions. In pelvic imaging, this mechanism is vital for preserving the sharp boundaries of the pelvic girdle and soft-tissue interfaces that are otherwise smoothed out in a standard LDM bottleneck.

\textbf{LoRA Adaptation:} 
The model employs Low-Rank Adaptation (LoRA) to fine-tune the UNet and VAE specifically for the pelvic modality translation task. By targeting key attention modules and convolutional kernels, LoRA enables the model to learn the specific Hounsfield-to-Intensity mapping of pelvic tissues while retaining the strong spatial priors of the base diffusion model.

\textbf{Conditioning Mechanism (Cross-Attention):} 
The CT conditioning is integrated through spatial cross-attention layers within the UNet. These layers allow the denoising process to be spatially guided by CT-derived features, ensuring that the generated MRI is anatomically registered to the source pelvic scan.

\section{TRAINING WORKFLOW}
Training involves a forward diffusion process (adding noise) and a learned reverse process (denoising).

\begin{figure}[H]
\centering
\resizebox{0.8\textwidth}{!}{
\begin{tikzpicture}[
    node distance=8mm,
    box/.style={draw=brand, thick, fill=softbg, minimum width=2cm, minimum height=0.7cm, align=center, rounded corners=1mm, font=\scriptsize},
    arr/.style={-{Latex[length=1.5mm]}, thick, draw=accent}
]
    \node[box] (z0) {$z_0$ (Real)};
    \node[box, right=of z0] (zt) {$z_t$ (Noisy)};
    \node[box, right=of zt] (pred) {Pred $\hat{\epsilon}$};
    \node[box, below=1cm of zt] (unet) {UNet $\epsilon_\theta$};
    \node[box, left=of unet] (ct) {CT Condition};
    \node[box, right=of unet] (time) {Time $t$};

    \draw[arr] (z0) -- (zt) node[midway, above, font=\tiny] {$\alpha_t, \sigma_t$};
    \draw[arr] (zt) -- (unet);
    \draw[arr] (ct) -- (unet);
    \draw[arr] (time) -- (unet);
    \draw[arr] (unet) -- (pred);
    
    \node[draw=accent, thick, fill=accent!5, below=5mm of pred, font=\tiny\bfseries] (loss) {Loss: $\|\epsilon - \hat{\epsilon}\|^2$};
    \draw[arr] (pred) -- (loss);
\end{tikzpicture}
}
\caption{Forward and Reverse Diffusion: Learning to denoise conditioned on CT and Time.}
\end{figure}

\textbf{Objective Function and Sampler Dynamics:} 
The model minimizes the Mean Squared Error (MSE) between the predicted noise and the actual noise added to the latents. At inference time, we use a \texttt{DDPMScheduler} to perform the reverse diffusion process. The sampler is configured to produce a deterministic reconstruction, ensuring that clinical interpretations remain consistent across repeated forward passes of the same CT volume.

\section{INFERENCE PIPELINE}
Inference (sampling) is an iterative process where the model starts from random noise and gradually removes it to reveal the synthetic MRI.

\begin{figure}[H]
\centering
\resizebox{0.9\textwidth}{!}{
\begin{tikzpicture}[
    node distance=8mm,
    block/.style={draw=brand, thick, fill=softbg, minimum width=2.4cm, minimum height=0.8cm, align=center, rounded corners=1mm, font=\scriptsize},
    arr/.style={-{Latex[length=1.5mm]}, thick, draw=brand!70}
]
    \node[block] (noise) {Gaussian Noise\\$z_T \sim \mathcal{N}(0, I)$};
    \node[block, right=of noise] (loop) {Iterative Denoising\\(50-100 steps)};
    \node[block, right=of loop] (z0) {Latent $z_0$};
    \node[block, right=of z0] (dec) {VAE Decoder};
    \node[block, below=1cm of dec] (out) {Synthetic MRI};

    \draw[arr] (noise) -- (loop);
    \draw[arr] (loop) -- (z0);
    \draw[arr] (z0) -- (dec);
    \draw[arr] (dec) -- (out);

    \node[block, above=8mm of loop] (ct) {CT Condition};
    \draw[arr] (ct) -- (loop);
\end{tikzpicture}
}
\caption{Inference Sampling: Iterative reconstruction guided by CT conditioning.}
\end{figure}

During inference, we use schedulers like DDIM or PNDM to accelerate the sampling process. The input CT volume is encoded once, and its features are used as constant conditioning throughout the denoising iterations. This ensures that every generated MRI slice is perfectly aligned with the anatomical structures present in the CT, which is critical for radiotherapy planning in the pelvis region.

\end{document}
